Quantitative Benchmarking for Theory Exploration

### 2017-08-07 TEBenchmarkPaper/article.tex
Added meeting notes

\author{Chris Warburton% \and
                %Second Author % etc
}

@@

\maketitle

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
The field of mathematics has been transformed by the introduction of computers:
calculations are now routinely automated, allowing humans to focus their efforts
on higher-level, more creative activities like proving theorems; \emph{automated
theorem proving} (ATP) seeks to automate this process too, and was one of the
first tasks studied in the field of artificial intelligence
\cite{newell1956logic,sutcliffe2001evaluating}. Despite this effort, ATP
methods are not widely used by traditional mathematicians, although they have
seen use in applied areas such as hardware and software verification
\cite{Moore:2003}.

@@

One problem with this approach is the small size of these libraries. For
example, a benchmark based on Isabelle/HOL's theory of natural numbers contains
only FIXME definitions and FIXME theorems. Whilst such benchmarks allow
comparison between different approaches, their narrow scope doesn't provide much
indication of performance in different, especially \emph{novel}, domains.

@@

%\includegraphics{bucketing-graph}

\begin{acknowledgements}
We are grateful for Jianguo Zhang
\end{acknowledgements}

>>>>

\author{Chris Warburton% \and
        Alison Pease
        % TODO: Invite Katya? Even if she only offers some help, can acknowledge
}

@@

\maketitle
IGNORE ISABELLE THEORIES FOR NOW!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
IGNORE CLUSTERING FOR NOW!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

intro (motivation, high-level background, eval is a problem),
background,
our proposal,
application (evaluation),
discussion (e.g. philosophical bits),
future work, conclusions, etc.

COMPARE COLTON ET AL.

Evaluation of conjectures, sets of conjectures, systems
what constitutes a good/desired output (in each author's terms)? This combines
their evaluation with their (system-specific) assumptions (e.g. IsaCoSy doesn't
generate redundant terms: non-redundancy is clearly a goal, but it's not measured
in eval since it's assumed based on system's construction)

Look carefully at each system's (paper's) eval section
Make a table like tabl 1 of Colton et al

Make it clear that methodology is the point (it's not just a detail, like in a
chemistry paper for example)

\begin{abstract}
  % TODO: Write abstract sooner rather than layer
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}

% TODO: Preaching to the choir;
% Jumping around a bit, rephrase
% Look at phd symposium 2017, moa's conjecture synthesis paper, etc.
% We can just jump straight in to automated conjecture generation

% Maybe lay out argument in bullet points; flesh out sentences later

% Lay out the field of TE: motivation, systems, evaluation, problems...
% Compare to

The field of mathematics has been transformed by the introduction of computers:
calculations are now routinely automated, allowing humans to focus their efforts
on higher-level, more creative activities like proving theorems; \emph{automated
theorem proving} (ATP) seeks to automate this process too, and was one of the
first tasks studied in the field of artificial intelligence
\cite{newell1956logic,sutcliffe2001evaluating}. Despite this effort, ATP
methods are not widely used by traditional mathematicians, although they have
seen use in applied areas such as hardware and software verification
\cite{Moore:2003}.

% FIXME: Be more specific about our contributions

@@

One problem with this approach is the small size of these libraries. For
example, a benchmark based on Isabelle/HOL's theory of natural numbers contains
only 4 definitions and 12 theorems. Whilst such benchmarks allow
comparison between different approaches, their narrow scope doesn't provide much
indication of performance in different, especially \emph{novel}, domains.

@@

We have applied our benchmarking methodology to the QuickSpec and IsaCoSy theory
exploration systems. To assess the scalability of their algorithms, we sampled
theories of various sizes: from theories containing only a single definition, up
to theories containing 20 definitions. These are shown in figure FIXME.

To provide a more robust comparison of these two systems, we also applied a
paired difference test: measuring, for each sampled theory, the difference in
time, precision and recall between the two systems.

\section{Conclusion}
\label{sec:conclusion}

Our benchmark suite provides a unifying goal for the diverse approaches of
theory exploration, whilst avoiding some of the philosophical complications of
the field. Whilst the current implementation is quite limited, we welcome
additions from other researchers, to more closely align the benchmark with their
goals, and the strenghts and weaknesses of their systems.

``Solving'' this benchmark suite would not solve the problem of theory exploration in general,
so more ambitious goals must be set in the future; but we have found that there is still a
long way to go until that problem arises.

\begin{acknowledgements}
  We are grateful to Jianguo Zhang for help with our statistical analysis.

  EPSRC
\end{acknowledgements}
